Simulation Tools for Small Area Estimation: Introducing the \proglang{R}-package \proglang{saeSim}

### 2014-10-20 saeSim.Rnw
raw case study

\section{Small area estimation}
\label{sec:SAE}
The objective of small area estimation is to produce reliable statistics (means, quantiles, proportions, etc.) for domains where little or no sampled units are available. Groups may be areas or other entities, for example defined by socio-economic characteristics. The demand for such estimators is rising as they are used for fund allocation, educational and health programs \citep{pfeffermann13}. As direct estimation of such statistics are considered to be unreliable, methods in small area estimation try to improve the domain predictions by borrowing strength from neighboured or \textit{similar} domains. This can be achieved by using additional information from census data to assist the prediction for non-sampled domains or domains with little information. For the purpose of this article we will introduce two basic models frequently used in small area estimation, the unit-level model introduced by \cite{battese1988} and the area-level model introduced by \cite{fay79}.

@@

With \proglang{saeSim} we want to contribute tools for simulation studies in the field of small area estimation. We see the need for sharing tools for data generation and simulation amongst the scientific community and thus defined an interface for these tools as well as designed a platform to make them accessible. By defining the steps behind a simulation we hope to promote a reasonable way to communicate them alongside publications and during research. In the next section we will present a case study and focus more on the concrete functionality provided by the package.

\section{Case study}
\label{sec:caseStudy}

\section{Outlook}
Use this package to share and publish simulation studies alongside papers. Contribute to the package to make your ideas available. Contribute to the package and make your whole simulation study available.

>>>>

\section{Small area estimation}
\label{sec:SAE}
The objective of small area estimation is to produce reliable statistics (means, quantiles, proportions, etc.) for domains where little or no sampled units are available. Groups may be areas or other entities, for example defined by socio-economic characteristics. The demand for such estimators is rising as they are used for fund allocation, educational and health programs \citep{pfeffermann13}. As direct estimation of such statistics are considered to be unreliable, methods in small area estimation try to improve the domain predictions by borrowing strength from neighboured or \textit{similar} domains. This can be achieved by using additional information from census data to assist the prediction for non-sampled domains or domains with little information. For the purpose of this article we will introduce two basic models frequently used in small area estimation, the unit-level model introduced by \cite{battese88} and the area-level model introduced by \cite{fay79}.

The unit level model \citep{battese88} can be expressed as:
\begin{eqnarray}
	 y_{\indexDomain\indexUnit} =& \xUnit^\top\beta + \randomEffectIndexed + \samplingErrorUnitIndexed \nonumber \\
	\randomEffectIndexed \stackrel{iid}{\sim}& N(0, \randomEffectVariance)  \nonumber \\
	\samplingErrorUnitIndexed \stackrel{iid}{\sim}& N(0, \samplingVariance) \nonumber
\end{eqnarray}

where $\indexDomain = 1, \dots, \nDomains$ and $\indexUnit = 1, \dots, \nUnitIndexed$. $y_{\indexDomain\indexUnit}$ is the the dependent variable for domain $\indexDomain$ and unit $\indexUnit$, and $\xUnit$ the corresponding auxiliary information for that unit. Furthermore $\randomEffect$ and $\samplingError$ are independent. This model can be seen as a linear mixed model from which the best linear unbiased predictor (BLUP) can be derived and is used for the domain prediction \citep{rao03}.

@@

With \proglang{saeSim} we want to contribute tools for simulation studies in the field of small area estimation. We see the need for sharing tools for data generation and simulation amongst the scientific community and thus defined an interface for these tools as well as a platform to make them accessible. By defining the steps behind a simulation we hope to promote a reasonable way to communicate them alongside publications and during research. In the next section we will present a case study and focus more on the concrete functionality provided by the package.

\section{Case study}
\label{sec:caseStudy}
\subsection{Model-based simulation}
<<>>=
suppressPackageStartupMessages(library(saeSim))
suppressPackageStartupMessages(library(sae))
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(reshape2))
suppressPackageStartupMessages(library(ggplot2))

set.seed(1) # for reproduction of the results
@

<<>>=
comp_FH <- function(dat) {
  # Computes the EBLUP under a Fay-Herriot model
  modelFH <- eblupFH(y ~ x, vardir, data = dat, method = "FH")
  dat$FH <- modelFH$eblup
  dat
}
@

<<>>=
simResults <- base_id(nDomains = 40, nUnits = 1) %>% 
  sim_gen_fe() %>%
  sim_gen_re() %>% 
  sim_gen(comp_var(vardir = seq(0.1, 4, length.out = 40))) %>%
  sim_gen(comp_var(e = rnorm(nrow(dat), sd = sqrt(vardir)))) %>%
  sim_resp_eq(y = 100 + 2 * x + v + e) %>%
  sim_comp_agg(comp_var(trueStat = y - e)) %>%
  sim_comp_agg(comp_FH) %>%
  sim(R = 500) %>%
  rbind_all
@

<<>>=
ggDat <- simResults %>% 
  melt(id.vars = c("idD", "trueStat"), 
       measure.vars = c("FH", "y"), 
       variable.name = "method",
       value.name = "prediction") %>% 
  mutate(RBIAS = (prediction - trueStat)/trueStat,
         RRMSE = (prediction - trueStat)^2/trueStat) %>%
  group_by(idD, method) %>% 
  summarise(RBIAS = mean(RBIAS),
            RMSE = mean(RRMSE)) %>%
  as.data.frame

head(ggDat)
@

<<modelSimBIAS, fig=TRUE, width = 8, height=4, include=FALSE>>=
ggplot(ggDat, aes(y = RBIAS, x = method)) + geom_boxplot() + coord_flip()
@

\includegraphics[width = \textwidth]{saeSim-modelSimBIAS.pdf}

<<modelSimRMSE, fig=TRUE, width = 8, height=4, include=FALSE>>=
ggplot(ggDat, aes(y = RMSE, x = method)) + geom_boxplot() + coord_flip()
@

\includegraphics[width = \textwidth]{saeSim-modelSimRMSE.pdf}

\subsection{Design-based simulation}
<<>>=
data(eusilcP, package = "simFrame")

simDat <- eusilcP %>% 
  mutate(agesq = age^2, eqIncome = as.numeric(eqIncome)) %>%
  filter(main) %>%
  select(region, eqIncome, age, agesq, gender) %>%
  base_add_id(c("region"))

attr(simDat, "popMeans") <- 
  simDat %>% 
  group_by(idD, region) %>%
  summarise(age = mean(age),
            agesq = mean(agesq),
            genderFemale = mean(as.integer(gender) - 1),
            trueStat = mean(eqIncome))

attr(simDat, "popN") <- 
  simDat %>%
  group_by(idD) %>%
  summarise(N = n())

attr(simDat, "popN")
attr(simDat, "popMeans")

@

<<>>=
comp_BHF <- function(dat) {
  popMeans <- attr(dat, "popMeans") %>% select(-trueStat, -region)
  modelBHF <- 
    eblupBHF(eqIncome ~ age + agesq + gender, idD, meanxpop = popMeans,
             popnsize = attr(dat, "popN"), data = dat)
  attr(dat, "BHF") <- modelBHF$eblup
  dat
}

comp_direct <- function(dat) {
  attr(dat, "sampleMean") <- 
    dat %>% group_by(idD) %>% summarise(y = mean(eqIncome))
  dat  
}

agg_results <- function(dat) {
  cbind(attr(dat, "popMeans") %>% select(idD, region, trueStat), 
        BHF = attr(dat, "BHF")$eblup,
        y = attr(dat, "sampleMean")$y)
}
@

<<>>=
simResults <- simDat %>% 
  sim_sample(sample_fraction(0.1, groupVars = "idD")) %>%
  sim_comp_sample(comp_BHF) %>%
  sim_comp_sample(comp_direct) %>%
  sim_agg(agg_results) %>%
  sim(R = 50) %>%
  rbind_all

ggDat <- simResults %>% 
  melt(id.vars = c("idD", "region", "trueStat"), 
       measure.vars = c("BHF", "y"), 
       variable.name = "method",
       value.name = "prediction")
@

<<designSimBIAS, fig=TRUE, width = 8.3, height=4, include=FALSE>>=
ggplot(ggDat, aes(x = region, y = prediction - trueStat, fill = method)) + 
  geom_boxplot() + theme(legend.position = "bottom")
@

\includegraphics[width = \textwidth]{saeSim-designSimBIAS.pdf}

<<designSimMSE, fig=TRUE, width = 8.3, height=4, include=FALSE>>=
ggplot(ggDat, aes(x = region, y = (prediction - trueStat)^2, fill = method)) + 
  geom_boxplot() + scale_y_log10() + theme(legend.position = "bottom")
@

\includegraphics[width = \textwidth]{saeSim-designSimMSE.pdf}

\section{Outlook}
Use this package to share and publish simulation studies alongside papers. Contribute to the package to make your ideas available. Contribute to the package and make your whole simulation study available. 
\begin{itemize}
	\item Link to simFrame
	\item Package Features
	\begin{itemize}
		\item outliers
		\item sampling
		\item non-linear models
	\end{itemize}
	\item How to contribute?
	\item parallel computations
\end{itemize}

### 2014-10-20 saeSim.tex
raw case study

\begin{document}
%
%% include your article here, just as usual
%
% R CMD Sweave %.Rnw
% Loading (and installing) necessary packages.
%
\section{Introduction}
Reproducible Research has become a widely discussed topic in general and also the field of statistics. Thanks to the many mostly open-source tools like the \proglang{R}-language \citep{r14} and \LaTeX, and also packages like knitr \citep{yihui13} and Sweave \citep{leisch02} and more recently rmarkdown \citep{allaire14}, the integration text and source code for statistical analysis is possible. Not only are tools available to make research reproducible, also the demand of making the analysis of articles reproducible is rising. Publishing source-code and data alongside an article draws special attention to authoring an analysis. However, the requirements in style and clarity of source code are different from the written words in the article itself.

@@

Small Area Estimation is a growing field inside the field of statistics, where simulation studies play an important role. New statistical models are applied in model-based and design-based simulation studies. Considering the demands in reproducibility we want to propose a framework for simulation studies. This framework addresses three demands. First, making tools for data generation available and reusable. Second, unify the process behind simulation studies inside the field of Small Area Estimation. And third, making source-code of simulation studies available, such that it supports the conducted research in a transparent manner.

In this article we want to introduce a new package for the \proglang{R}-language addressing these demands. We will show the importance of simulation studies in the field, introduce the simulation framework and demonstrate how to use the package to map the simulation as a process to \proglang{R}.

\section{Small area estimation}
The objective of small area estimation is to produce reliable statistics (means, quantiles, proportions, etc.) for domains where little or no sampled units are available. Groups may be areas or other entities, for example defined by socio-economic characteristics. The demand for such estimators is rising as they are used for fund allocation, educational and health programs \citep{pfeffermann13}. As direct estimation of such statistics are considered to be unreliable, methods in small area estimation try to improve the domain predictions by borrowing strength from neighboured or \textit{similar} domains. This can be achieved by using additional information from census data to assist the prediction for non-sampled or domains with little information. For the purpose of this article we will introduce two basic models frequently used in small area estimation, the unit-level model introduced by \cite{battese1988} and the area-level model introduced by \cite{fay79}.

As applications in small area estimations rely often on sensitive information unit-level information is not always available. Instead only aggregates, or rather the direct estimators are. In such situations area-level models are used. The area-level model introduced by \cite{fay79} is build on a sampling model:
%
\[\directStat_{\indexDomain} = \trueStat_{\indexDomain} + \samplingError_{\indexDomain},\]
%
where $\directStat_{\indexDomain}$ is a direct estimator of a statistic of interest $\trueStat_{\indexDomain}$ for an area $\indexDomain$ with $\indexDomain = 1, \dots, \nDomains$ and $\nDomains$ being the total number of areas. The sampling error $\samplingError_{\indexDomain}$ is assumed to be independent and normally distributed with known variances $\samplingVarianceIndexed$, i.e. $\samplingError_{\indexDomain}|\trueStat_{\indexDomain} \sim \mathit{N}(0, \samplingVarianceIndexed)$. The model is modified with the linking model by assuming a linear relationship between the true area statistic $\trueStat_{\indexDomain}$ and some auxiliary variables $\xArea$:
%
\[\trueStat_{\indexDomain} = \xArea^\top \beta + \randomEffectIndexed,\] 
%
with $\indexDomain=1,\dots, \nDomains$. Note that $\xArea$ is a vector containing area-level (aggregated) information for $\nRegressor$ variables and $\beta$ is a vector ($1\times \nRegressor$) of regression coefficients describing the (linear) relationship. The model errors $\randomEffectIndexed$ are assumed to be independent and normally distributed, i.e. $\randomEffectIndexed \sim \mathit{N}(0, \randomEffectVariance)$. Furthermore $\samplingErrorIndexed$ and $\randomEffectIndexed$ are assumed to be independent. Combining the sampling and linking model leads to:
\begin{equation}
\label{eq:FH}
\directStatIndexed = \xArea^\top \beta + \randomEffectIndexed + \samplingErrorIndexed.
\end{equation} 
%
Model \ref{eq:FH} is effectively a random-intercept model where the distribution of the error term $\samplingErrorIndexed$ is heterogeneous and known.

\section{A simulation framework}
%
%% flow-diagram
\begin{wrapfigure}{R}{0.5\textwidth}
\begin{center}
\includegraphics[width=0.5\textwidth]{flowdiagram}
\end{center}
\caption{\label{fig:flowdiagram}Process of simulation.} 
\end{wrapfigure}
%
In our opinion simulation studies can best be summarised when understood as a process of data manipulation. Thus the main focus of \proglang{saeSim} is to define the steps in that process which can then be \textit{cleanly} defined and repeated. Before we go into any detail of the functionality of the package we will discuss the process behind simulation studies and later how \proglang{saeSim} maps this process into \proglang{R}.

@@

With \proglang{saeSim} we want to contribute tools for simulation studies in the field of small area estimation. We see the need for sharing tools for data generation and simulation amongst the scientific community and thus defined an interface for these tools as well as designed a platform to make them accessible. By defining the steps behind a simulation we hope to promote a reasonable way to communicate them alongside publications and during research. In the next section we will present a case study and focus more on the concrete functionality provided by the package.

\section{Case study}

\section{Outlook}
Use this package to share and publish simulation studies alongside papers. Contribute to the package to make your ideas available. Contribute to the package and make your whole simulation study available.

>>>>

\begin{document}
%
%% include your article here, just as usual
%
% R CMD Sweave %.Rnw
% Loading (and installing) necessary packages.
%
\section{Introduction}
The objective of small area estimation is to produce reliable statistics (means, quantiles, proportions, etc.) for domains where little or no sampled units are available. Groups may be areas or other entities, for example defined by socio-economic characteristics. New statistical methods are applied in model-based and design-based simulation studies. Considering the demands for reproducible research we propose a framework for simulation studies inside the field of small area estimation.

Reproducible Research has become a widely discussed topic in general and also in the field of statistics. Thanks to the many mostly open-source tools like the \proglang{R}-language \citep{r14} and \LaTeX, and also packages like knitr \citep{yihui13} and Sweave \citep{leisch02} and more recently rmarkdown \citep{allaire14}, the integration of text and source code for statistical analysis is possible. Not only are tools available to make research reproducible, also the demand of making the analysis of articles reproducible is rising which is supported by the rising number of Open Access Journals. Publishing source-code and data alongside an article draws special attention to authoring an analysis. However, the requirements for source code are different from the written words in the article itself.

@@

In this article we want to introduce a new package for the \proglang{R}-language to support the process of making simulation studies reproducible as well as supporting a \textit{human-readable} interface. With this package we have three objectives in specific: First, making tools for data generation available and reusable. Second, unify the process behind simulation studies inside the field of Small Area Estimation. And third, making source-code of simulation studies available, such that it supports the conducted research in a transparent manner.

In section \ref{sec:SAE} we will give a short introduction to small area estimation and the role of simulation studies within that field. Section \ref{sec:framework} will then introduce a framework for simulation studies and how it is supported by \proglang{saeSim}. To illustrate some of the package's features we present a case study in section \ref{sec:caseStudy}.

\section{Small area estimation}
\label{sec:SAE}
The objective of small area estimation is to produce reliable statistics (means, quantiles, proportions, etc.) for domains where little or no sampled units are available. Groups may be areas or other entities, for example defined by socio-economic characteristics. The demand for such estimators is rising as they are used for fund allocation, educational and health programs \citep{pfeffermann13}. As direct estimation of such statistics are considered to be unreliable, methods in small area estimation try to improve the domain predictions by borrowing strength from neighboured or \textit{similar} domains. This can be achieved by using additional information from census data to assist the prediction for non-sampled domains or domains with little information. For the purpose of this article we will introduce two basic models frequently used in small area estimation, the unit-level model introduced by \cite{battese88} and the area-level model introduced by \cite{fay79}.

The unit level model \citep{battese88} can be expressed as:
\begin{eqnarray}
	 y_{\indexDomain\indexUnit} =& \xUnit^\top\beta + \randomEffectIndexed + \samplingErrorUnitIndexed \nonumber \\
	\randomEffectIndexed \stackrel{iid}{\sim}& N(0, \randomEffectVariance)  \nonumber \\
	\samplingErrorUnitIndexed \stackrel{iid}{\sim}& N(0, \samplingVariance) \nonumber
\end{eqnarray}

where $\indexDomain = 1, \dots, \nDomains$ and $\indexUnit = 1, \dots, \nUnitIndexed$. $y_{\indexDomain\indexUnit}$ is the the dependent variable for domain $\indexDomain$ and unit $\indexUnit$, and $\xUnit$ the corresponding auxiliary information for that unit. Furthermore $\randomEffect$ and $\samplingError$ are independent. This model can be seen as a linear mixed model from which the best linear unbiased predictor (BLUP) can be derived and is used for the domain prediction \citep{rao03}.

Due to reasons of confidentiality unit-level information is not always available. Instead only aggregates, or rather the direct estimators may be supplied. However, these direct estimations are known to be unreliable, hence in such situations area-level models can be valuable. The area-level model introduced by \cite{fay79} is build on a sampling model:
%
\[\directStat_{\indexDomain} = \trueStat_{\indexDomain} + \samplingError_{\indexDomain},\]
%
where $\directStat_{\indexDomain}$ is a direct estimator of a statistic of interest $\trueStat_{\indexDomain}$ for an area $\indexDomain$ with $\indexDomain = 1, \dots, \nDomains$ and $\nDomains$ being the total number of areas. The sampling error $\samplingError_{\indexDomain}$ is assumed to be independent and normally distributed with known variances $\samplingVarianceIndexed$, i.e. $\samplingError_{\indexDomain}|\trueStat_{\indexDomain} \sim \mathit{N}(0, \samplingVarianceIndexed)$. The model is modified with the linking model by assuming a linear relationship between the true area statistic $\trueStat_{\indexDomain}$ and some auxiliary variables $\xArea$:
%
\[\trueStat_{\indexDomain} = \xArea^\top \beta + \randomEffectIndexed,\] 
%
with $\indexDomain=1,\dots, \nDomains$. Note that $\xArea$ is a vector containing area-level (aggregated) information for $\nRegressor$ variables and $\beta$ is a vector ($1\times \nRegressor$) of regression coefficients describing the (linear) relationship. The model errors $\randomEffectIndexed$ are assumed to be independent and normally distributed, i.e. $\randomEffectIndexed \sim \mathit{N}(0, \randomEffectVariance)$. Furthermore $\samplingErrorIndexed$ and $\randomEffectIndexed$ are assumed to be independent. Combining the sampling and linking model leads to:
\begin{equation}
\label{eq:FH}
\directStatIndexed = \xArea^\top \beta + \randomEffectIndexed + \samplingErrorIndexed.
\end{equation} 
%
Model \ref{eq:FH} is effectively a random-intercept model where the distribution of the error term $\samplingErrorIndexed$ is heterogeneous and known.

\section{A simulation framework}
\label{sec:framework}
%
%% flow-diagram
\begin{wrapfigure}{R}{0.5\textwidth}
\begin{center}
\includegraphics[width=0.5\textwidth]{flowdiagram}
\end{center}
\caption{\label{fig:flowdiagram}Process of simulation.} 
\end{wrapfigure}
%
In our opinion simulation studies can best be summarised when understood as a process of data manipulation. Thus the main focus of \proglang{saeSim} is to define the steps in that process which can then be \textit{cleanly} defined and repeated. Before we go into any detail of the functionality of the package we will discuss the process behind simulation studies and later how \proglang{saeSim} maps this process into \proglang{R}.

@@

With \proglang{saeSim} we want to contribute tools for simulation studies in the field of small area estimation. We see the need for sharing tools for data generation and simulation amongst the scientific community and thus defined an interface for these tools as well as a platform to make them accessible. By defining the steps behind a simulation we hope to promote a reasonable way to communicate them alongside publications and during research. In the next section we will present a case study and focus more on the concrete functionality provided by the package.

\section{Case study}
\label{sec:caseStudy}
\subsection{Model-based simulation}
\begin{Schunk}
\begin{Sinput}
> suppressPackageStartupMessages(library(saeSim))
> suppressPackageStartupMessages(library(sae))
> suppressPackageStartupMessages(library(dplyr))
> suppressPackageStartupMessages(library(reshape2))
> suppressPackageStartupMessages(library(ggplot2))
> set.seed(1) # for reproduction of the results
\end{Sinput}
\end{Schunk}

\begin{Schunk}
\begin{Sinput}
> comp_FH <- function(dat) {
+   # Computes the EBLUP under a Fay-Herriot model
+   modelFH <- eblupFH(y ~ x, vardir, data = dat, method = "FH")
+   dat$FH <- modelFH$eblup
+   dat
+ }
\end{Sinput}
\end{Schunk}

\begin{Schunk}
\begin{Sinput}
> simResults <- base_id(nDomains = 40, nUnits = 1) %>% 
+   sim_gen_fe() %>%
+   sim_gen_re() %>% 
+   sim_gen(comp_var(vardir = seq(0.1, 4, length.out = 40))) %>%
+   sim_gen(comp_var(e = rnorm(nrow(dat), sd = sqrt(vardir)))) %>%
+   sim_resp_eq(y = 100 + 2 * x + v + e) %>%
+   sim_comp_agg(comp_var(trueStat = y - e)) %>%
+   sim_comp_agg(comp_FH) %>%
+   sim(R = 500) %>%
+   rbind_all
\end{Sinput}
\end{Schunk}

\begin{Schunk}
\begin{Sinput}
> ggDat <- simResults %>% 
+   melt(id.vars = c("idD", "trueStat"), 
+        measure.vars = c("FH", "y"), 
+        variable.name = "method",
+        value.name = "prediction") %>% 
+   mutate(RBIAS = (prediction - trueStat)/trueStat,
+          RRMSE = (prediction - trueStat)^2/trueStat) %>%
+   group_by(idD, method) %>% 
+   summarise(RBIAS = mean(RBIAS),
+             RMSE = mean(RRMSE)) %>%
+   as.data.frame
> head(ggDat)
\end{Sinput}
\begin{Soutput}
  idD method         RBIAS         RMSE
1   1     FH -0.0001326439 0.0009412468
2   1      y -0.0001285281 0.0009413428
3   2     FH  0.0002999555 0.0017042498
4   2      y  0.0002214372 0.0018404586
5   3     FH  0.0004056162 0.0027148056
6   3      y  0.0003529474 0.0032745696
\end{Soutput}
\end{Schunk}

\begin{Schunk}
\begin{Sinput}
> ggplot(ggDat, aes(y = RBIAS, x = method)) + geom_boxplot() + coord_flip()
\end{Sinput}
\end{Schunk}

\includegraphics[width = \textwidth]{saeSim-modelSimBIAS.pdf}

\begin{Schunk}
\begin{Sinput}
> ggplot(ggDat, aes(y = RMSE, x = method)) + geom_boxplot() + coord_flip()
\end{Sinput}
\end{Schunk}

\includegraphics[width = \textwidth]{saeSim-modelSimRMSE.pdf}

\subsection{Design-based simulation}
\begin{Schunk}
\begin{Sinput}
> data(eusilcP, package = "simFrame")
> simDat <- eusilcP %>% 
+   mutate(agesq = age^2, eqIncome = as.numeric(eqIncome)) %>%
+   filter(main) %>%
+   select(region, eqIncome, age, agesq, gender) %>%
+   base_add_id(c("region"))
> attr(simDat, "popMeans") <- 
+   simDat %>% 
+   group_by(idD, region) %>%
+   summarise(age = mean(age),
+             agesq = mean(agesq),
+             genderFemale = mean(as.integer(gender) - 1),
+             trueStat = mean(eqIncome))
> attr(simDat, "popN") <- 
+   simDat %>%
+   group_by(idD) %>%
+   summarise(N = n())
> attr(simDat, "popN")
\end{Sinput}
\begin{Soutput}
Source: local data frame [9 x 2]

  idD    N
1   1  799
2   2 4619
3   3 5857
4   4 1723
5   5 3386
6   6 4071
7   7 1671
8   8 1889
9   9  985
\end{Soutput}
\begin{Sinput}
> attr(simDat, "popMeans")
\end{Sinput}
\begin{Soutput}
Source: local data frame [9 x 6]
Groups: idD

  idD        region      age    agesq genderFemale trueStat
1   1    Burgenland 54.50063 3269.677    0.3366708 22005.42
2   2 Lower Austria 51.95259 3009.934    0.3777874 19813.37
3   3        Vienna 46.98310 2486.448    0.4662797 20395.84
4   4     Carinthia 51.81428 2995.735    0.3540337 19486.18
5   5        Styria 50.64087 2886.845    0.3573538 19335.39
6   6 Upper Austria 50.18644 2795.804    0.3443871 20517.29
7   7      Salzburg 51.44943 2965.268    0.4189108 19890.33
8   8         Tyrol 51.76707 2995.451    0.3975648 19350.89
9   9    Vorarlberg 49.06904 2697.382    0.3583756 22156.12
\end{Soutput}
\begin{Sinput}
> 
\end{Sinput}
\end{Schunk}

\begin{Schunk}
\begin{Sinput}
> comp_BHF <- function(dat) {
+   popMeans <- attr(dat, "popMeans") %>% select(-trueStat, -region)
+   modelBHF <- 
+     eblupBHF(eqIncome ~ age + agesq + gender, idD, meanxpop = popMeans,
+              popnsize = attr(dat, "popN"), data = dat)
+   attr(dat, "BHF") <- modelBHF$eblup
+   dat
+ }
> comp_direct <- function(dat) {
+   attr(dat, "sampleMean") <- 
+     dat %>% group_by(idD) %>% summarise(y = mean(eqIncome))
+   dat  
+ }
> agg_results <- function(dat) {
+   cbind(attr(dat, "popMeans") %>% select(idD, region, trueStat), 
+         BHF = attr(dat, "BHF")$eblup,
+         y = attr(dat, "sampleMean")$y)
+ }
\end{Sinput}
\end{Schunk}

\begin{Schunk}
\begin{Sinput}
> simResults <- simDat %>% 
+   sim_sample(sample_fraction(0.1, groupVars = "idD")) %>%
+   sim_comp_sample(comp_BHF) %>%
+   sim_comp_sample(comp_direct) %>%
+   sim_agg(agg_results) %>%
+   sim(R = 50) %>%
+   rbind_all
> ggDat <- simResults %>% 
+   melt(id.vars = c("idD", "region", "trueStat"), 
+        measure.vars = c("BHF", "y"), 
+        variable.name = "method",
+        value.name = "prediction")
\end{Sinput}
\end{Schunk}

\begin{Schunk}
\begin{Sinput}
> ggplot(ggDat, aes(x = region, y = prediction - trueStat, fill = method)) + 
+   geom_boxplot() + theme(legend.position = "bottom")
\end{Sinput}
\end{Schunk}

\includegraphics[width = \textwidth]{saeSim-designSimBIAS.pdf}

\begin{Schunk}
\begin{Sinput}
> ggplot(ggDat, aes(x = region, y = (prediction - trueStat)^2, fill = method)) + 
+   geom_boxplot() + scale_y_log10() + theme(legend.position = "bottom")
\end{Sinput}
\end{Schunk}

\includegraphics[width = \textwidth]{saeSim-designSimMSE.pdf}

\section{Outlook}
Use this package to share and publish simulation studies alongside papers. Contribute to the package to make your ideas available. Contribute to the package and make your whole simulation study available. 
\begin{itemize}
	\item Link to simFrame
	\item Package Features
	\begin{itemize}
		\item outliers
		\item sampling
		\item non-linear models
	\end{itemize}
	\item How to contribute?
	\item parallel computations
\end{itemize}
